---
title: "Proportion of population using of on-site sanitation facilities with human waste treated off-site"
subtitle: "Dhruv Sethi - Assignment 2"
author: "Damien Dupre"
format: html
date: 27/04/2023
execute:
  echo: false
  warning: false
  message: false
editor: visual
---

## The visualisation shown below is a world map

### GDP Per Capita

```{r}
# Load required packages
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(maps)
library(dplyr)


# Read in data files
data2 <- read.csv("data2.csv")
data3 <- read.csv("data3.csv")

# Combine data
combined_data <- left_join(data2, data3, by = c("country", "year"))
#print(combined_data)
#print(combined_data$GDP.per.capita)


# Create world map chart
world <- map_data("world")
ggplot() +
  geom_map(data = combined_data, map = world, aes(fill = GDP.per.capita, map_id = country), color = "black", size = 0.1) +
  expand_limits(x = world$long, y = world$lat) +
  labs(title = "World Map of GDP", fill = "GDP.per.capita")

```

### Obs Value

```{r}
# Load required packages
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(maps)
library(dplyr)


# Read in data files
data2 <- read.csv("data2.csv")
data3 <- read.csv("data3.csv")

# Combine data
combined_data <- left_join(data2, data3, by = c("country", "year"))
#print(combined_data)
#print(combined_data$GDP.per.capita)

world <- map_data("world")
ggplot() +
  geom_map(data = combined_data, map = world, aes(fill = obs_value, map_id = country), color = "green", size = 0.1) +
  expand_limits(x = world$long, y = world$lat) +
  labs(title = "World Map of obs_value", fill = "obs_value")
```

### ScatterPlot between GDP Per Capita and Obs Value in the South American countries

In terms of on-site sanitation specifically, the proportion of the population using improved on-site sanitation facilities which includes facilities that safely separate human waste from contact with people and the environment) increased from 46% in 1990 to 60% in 2015 in South America. It appears that while there has been progress in improving sanitation in South America, there is still room for improvement in terms of ensuring safe and sustainable sanitation for all.

```{r}
# Load required packages
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(maps)
library(dplyr)


# Read in data files
data2 <- read.csv("data2.csv")
data3 <- read.csv("data3.csv")

# Combine data
combined_data <- left_join(data2, data3, by = c("country", "year"))
#print(combined_data)
#print(combined_data$GDP.per.capita)

# Define South American countries
south_america <- c("Argentina", "Bolivia", "Brazil", "Chile", "Colombia", "Ecuador", "Guyana", "Paraguay", "Peru", "Suriname", "Venezuela")

# Filter data for South American countries
sa_data <- combined_data %>% filter(country %in% south_america)

# Create scatterplot with country colors
ggplot(sa_data, aes(x = GDP.per.capita, y = obs_value, fill = country)) +
  geom_point(shape = 21, size = 5) +
  scale_fill_manual(values = c(
    "Argentina" = "red",
    "Bolivia" = "blue",
    "Brazil" = "green",
    "Chile" = "orange",
    "Colombia" = "purple",
    "Ecuador" = "brown",
    "Guyana" = "pink",
    "Paraguay" = "gray",
    "Peru" = "black",
    "Suriname" = "violet",
    "Venezuela" = "cyan"
  )) +
  labs(title = "GDP.per.capita vs obs_value in South America", x = "GDP.per.capita", y = "obs_value")
```

### **Bar Chart for GDP Per Capita and South American Countries**

```{r}
# Load required packages
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(maps)
library(dplyr)


# Read in data files
data2 <- read.csv("data2.csv")
data3 <- read.csv("data3.csv")

# Combine data
combined_data <- left_join(data2, data3, by = c("country", "year"))
#print(combined_data)
#print(combined_data$GDP.per.capita)

# Create bar chart for South American countries
ggplot(sa_data, aes(x = country, y = GDP.per.capita)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "GDP by Country in South America", x = "country", y = "GDP.per.capita")
```

### **Time Series of both GDP Per Capita and Obs Value over the time period**

```{r}
# Load required packages
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(maps)
library(dplyr)


# Read in data files
data2 <- read.csv("data2.csv")
data3 <- read.csv("data3.csv")

# Combine data
combined_data <- left_join(data2, data3, by = c("country", "year"))
#print(combined_data)
#print(combined_data$GDP.per.capita)

# Create cleaned data for South American countries
sa_data_clean <- sa_data %>%
  gather(key = "indicator", value = "obs_value", -country, -year)

# Filter the data for GDP.per.capita and obs_value indicators
sa_data_filtered <- sa_data_clean %>%
  filter(indicator %in% c("GDP.per.capita", "obs_value"))

# Create time-series chart for South American countries
ggplot(sa_data_filtered, aes(x = year, y = obs_value, group = country, color = country)) +
  geom_line(size = 1.2) + # Increase line thickness for better visibility
  scale_color_brewer(type = "qual", palette = "Set2") + # Use a color palette with higher contrast
  facet_grid(rows = vars(indicator), cols = vars(country), scales = "free_y",
             switch = "y", space = "free") + # Add facets for each indicator and country
  labs(title = "Trend in Indicators over Time in South America", x = "", y = "") +
  theme_minimal() + # Use a minimal theme
  theme(legend.position = "bottom", legend.direction = "horizontal",
        strip.background = element_blank(), # Remove background color from facet labels
        panel.grid.major.y = element_line(color = "gray90"), # Add horizontal grid lines
        panel.spacing = unit(0.5, "lines"),
        axis.text.x = element_blank(), # Remove x-axis tick labels
        axis.ticks.x = element_blank(), # Remove x-axis tick marks
        axis.title.x = element_text(margin = margin(t = 20))) # Add x-axis title with margin
```

If the GDP per capita of a country is going up (meaning that the country's economic output per person is increasing), but the proportion of the population using on-site sanitation facilities with human waste treated off-site is going down, it could mean that the government and private sector are not prioritizing investments in sanitation infrastructure.

And, if the country's GDP per capita of a country is going up and the proportion of the population using on-site sanitation facilities with human waste treated off-site is also going up, this could mean that the country's government is investing in their sanitation infrastructure.

For the countries in South America, Argentina, Brazil and Ecudor, GDP is projecting an upward trend whereas, the percentage values are projecting a downward trend, telling that the government in these countries are less invested in thier sanitation infrastructure. On the other hand, countries like, Bolivia, Chile, Colombia and Peru are showing an upward trend in both the cases.

---
title: "Proportion of population using of on-site sanitation facilities with human waste treated off-site"
subtitle: "Dhruv Sethi - Assignment 2"
authoor: "Damien Dupre"
format: html
date: 27/04/2023
---
